Analyzing Forest Elephant Trails in Gabon, Africa

### 2021-04-27 Code/Project_Template.Rmd
Updates to project template

---
title: "draft final Project"
author: "Elise Harrigan and Katie Krejsa"
date: "4/19/2021"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(plyr)
library(stringr)
library(tidyr)
library(lubridate)
library(tidyverse)
library(agricolae) 
library(dplyr)
library(viridis)
library(RColorBrewer)
library(colormap)
library(sf)
library(leaflet)
library(mapview)
library(raster)
library(sp)
library(rgdal)
library(cowplot)
#install.packages("PMCMRplus")
library(PMCMRplus)
#install.packages("rnaturalearth")
#install.packages("rnaturalearthdata")
library(rnaturalearth)

#set theme
mytheme <- theme_light(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "bottom", legend.justification = "right")
theme_set(mytheme)

##load in data - change this to pull from the processed file
tracklogs <- read.csv("./Data/Processed/IV_tracklogs_processed.csv")
wpts <- read.csv("./Data/Processed/IV_wpts_processed.csv")

#change to data 
tracklogs$date <- as.Date(tracklogs$date, format = "%Y-%m-%d")
wpts$date <- as.Date(wpts$date, format = "%Y-%m-%d")
```

[knitting](../ENV872_EDA_FinalProject/img/elephant.png)

## ELEPHANT MOVEMENT IN GABON, AFRICA

@@

## STUDY QUESTIONS

@@

1. What is the condition and use of the trail? 
2. What is frequency and number of days traveled on each trail?

@@

## DATASETS

@@

__Data Wrangling__

@@

## EXPLORATORY ANALYSIS
The elephant trail dataset has dimensions of 40,910 observation and 11 variables. We looked primarily at the use and the width of each trail and the longitude and latitude. First, we wanted to see the total amount of trails for each use type (P, M, W) in order to see which type of trail was most common. By summarizing the total amounts of each use of trail, we were able to see the Poor trails were the most commonly described trail.

@@

```{r}
##Insert code for map
##subset for each type of trail for 04-15-2018
Poor_trails <- tracklogs %>%
  filter(use_cha == "P")

@@

#convert to sf and include crs
subset_sf <- st_as_sf(subset_trails, crs = 4326, coords = c("lon", "lat"))

subset_poor_sf <- st_as_sf(subset_poor, crs = 4326, coords = c("lon", "lat"))
subset_med_sf <- st_as_sf(subset_med, crs = 4326, coords = c("lon", "lat"))
subset_heavy_sf <- st_as_sf(subset_heavy, crs = 4326, coords = c("lon", "lat"))

@@

## ANALYSIS

@@

```{r}

@@

```{r plots by width}
##for caption: On the x axis, the three letters represent how heavily trodden the path is:  (P = poor, M = medium, W = heavy) and the width of the trail on the y-axis in centimeters. It is evident the poor and medium trails are similar in width and the heavy trails are much wider.

@@

```{r}
mod3 <- lm(data = tracklogs, use_num ~ use_cha)
  summary(mod3)
  TukeyHSD(aov(mod3))
  plot(TukeyHSD(aov(mod3)))
```

### SUMMARY AND CONCLUSIONS
In our final dataset, which included only the data that were recorded on common dates between our two initial datasets, there are 27,218 poor trodden trail points, 13,666 medium trodden trail points, and 26 heavily trodden trail points. From our analyses on the widths of each type of trail, we found that the heavily trodden trail is significantly wider than the poor and medium trodden trails and the widths of the poor and medium trodden trails are not significantly different. While the heavily trodden trails are significantly wider than the poor and medium trodden trails, it is important to note that the heavily trodden trails have a much smaller number of recorded trail points, which occur in a compact geographic area compared to all of the other recorded points, and only on one day. Overall, we can conclude that the majority of recorded trail points are poorly trodden (67%) and medium trodden (33%), and very few are heavily trodden (<0.1%).

The second question we wanted to answer was to see the number of trail points occurring for each day. In general, you can see that the level of use increased from poorly trodden to medium trodden over time. By looking at the graph, the poor trodden trails are more abundant earlier in the year and the medium trodden trails become more abundant in April and May.

Using these elephant trail data for the preliminary analyses we have presented in this project provides an initial exploration into the uses for these data. In future studies, these data could be used to see how the trails change due to human disturbance or presence or at how elephants change course or not. There are also opportunities with these data to look at the change of vegetation along the heavier trodden trails and if it will affect the reproduction, seed dispersal or the types of plants that come up in the opened spaces in the forest.

>>>>

---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    fig_height: 3  
    fig_width: 4
    number_sections: yes
geometry: margin=2.54cm
title: "Analyzing Forest Elephant Trails in Gabon, Africa"
subtitle: "https://github.com/kmk85/HarriganKrejsa_ENV872_EDA_FinalProject"
author: "Elise Harrigan and Katie Krejsa"
fontsize: 12pt
mainfont: Times New Roman
always_allow_html: true

---

\newpage
\tableofcontents 
\newpage
\listoffigures 
\newpage

```{r setup, echo=FALSE, include=FALSE}
# Set your working directory
getwd()

# Load your packages
library(plyr)
library(stringr)
library(tidyr)
library(lubridate)
library(tidyverse)
library(agricolae) 
library(dplyr)
library(viridis)
library(RColorBrewer)
library(colormap)
library(sf)
library(mapview)
library(raster)
library(sp)
library(rgdal)
library(cowplot)

# Set your ggplot theme
mytheme <- theme_light(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "bottom", legend.justification = "right")
theme_set(mytheme)

# Load your datasets
tracklogs <- read.csv("./Data/Processed/IV_tracklogs_processed.csv")
wpts <- read.csv("./Data/Processed/IV_wpts_processed.csv")

#change the dates
tracklogs$date <- as.Date(tracklogs$date, format = "%Y-%m-%d")
wpts$date <- as.Date(wpts$date, format = "%Y-%m-%d")

# Rationale and Research Questions

## Elephant Movement in Gabon, Africa

@@

## Study Questions

@@

1. What is frequency and number of days traveled on each trail? 
2. What is the condition and use of the trail?

@@

\newpage

# Dataset Information

@@

## Data Wrangling

@@

\newpage

# Exploratory Analysis

The elephant trail dataset has dimensions of 40,910 observation and 11 variables. We looked primarily at the use and the width of each trail and the longitude and latitude. First, we wanted to see the total amount of trails for each use type (P, M, W) in order to see which type of trail was most common. By summarizing the total amounts of each use of trail, we were able to see the Poor trails were the most commonly described trail.

@@

```{r, echo=FALSE, fig.cap="Map of all trail use types on April 15, 2018.", fig.align="left", warning=FALSE}
##Insert code for map
##subset for each type of trail for 04-15-2018
Poor_trails <- tracklogs %>%
  filter(use_cha == "P")

@@

#convert to sf and include crs
subset_poor_sf <- st_as_sf(subset_poor, crs = 4326, coords = c("lon", "lat"))
subset_med_sf <- st_as_sf(subset_med, crs = 4326, coords = c("lon", "lat"))
subset_heavy_sf <- st_as_sf(subset_heavy, crs = 4326, coords = c("lon", "lat"))

@@

\newpage

# Analysis

@@

```{r, echo=FALSE, fig.cap="The poor trodden trails are more abundant earlier in the year and the medium trodden trails take over in late April-May timeframe.", fig.align="left", warning=FALSE}

@@

```{r plots by width, echo=FALSE, fig.cap="On the x axis, the three letters represent how heavily trodden the path is:  (P = poor, M = medium, W = heavy) and the width of the trail on the y-axis in centimeters.", fig.align="left", warning=FALSE}
##for caption: On the x axis, the three letters represent how heavily trodden the path is:  (P = poor, M = medium, W = heavy) and the width of the trail on the y-axis in centimeters. It is evident the poor and medium trails are similar in width and the heavy trails are much wider.

@@

```{r, echo=FALSE, fig.cap="Post-hoc Tukey test for trail use.", fig.align="left", warning=FALSE}
mod3 <- lm(data = tracklogs, use_num ~ use_cha)
  plot(TukeyHSD(aov(mod3)))
```

\newpage

# Summary and Conclusions

The first question we wanted to answer was to see the number of trail points occurring for each day. In general, you can see that the level of use increased from poorly trodden to medium trodden over time. By looking at the graph, the poor trodden trails are more abundant earlier in the year and the medium trodden trails become more abundant in April and May.

In our final dataset, which included only the data that were recorded on common dates between our two initial datasets, there are 27,218 poor trodden trail points, 13,666 medium trodden trail points, and 26 heavily trodden trail points. From our analyses on the widths of each type of trail, we found that the heavily trodden trail is significantly wider than the poor and medium trodden trails and the widths of the poor and medium trodden trails are not significantly different. While the heavily trodden trails are significantly wider than the poor and medium trodden trails, it is important to note that the heavily trodden trails have a much smaller number of recorded trail points, which occur in a compact geographic area compared to all of the other recorded points, and only on one day. Overall, we can conclude that the majority of recorded trail points are poorly trodden (67%) and medium trodden (33%), and very few are heavily trodden (<0.1%).

The preliminary analyses presented in this project, provide an initial exploration into the uses for these data. In future studies, these data could be used to see how the trails change due to human disturbance or presence, or to see how and where elephants do or do not change course. There are also opportunities with these data to look at the change of vegetation along the different use trails and if it will affect the reproduction, seed dispersal or the types of plants that come up in the forest gaps.

### 2021-04-27 Code/Project_Template.tex
Updates to project template

---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    fig_height: 3  
    fig_width: 4
    number_sections: yes
geometry: margin=2.54cm
title: "Analyzing Forest Elephant Trails in Gabon, Africa"
subtitle: "https://github.com/kmk85/HarriganKrejsa_ENV872_EDA_FinalProject"
author: "Elise Harrigan and Katie Krejsa"
fontsize: 12pt
mainfont: Times New Roman
always_allow_html: true

---

@@

```{r setup, echo=FALSE, include=FALSE}
# Set your working directory
getwd()

# Load your packages
library(plyr)
library(stringr)
library(tidyr)
library(lubridate)
library(tidyverse)
library(agricolae) 
library(dplyr)
library(viridis)
library(RColorBrewer)
library(colormap)
library(sf)
library(mapview)
library(raster)
library(sp)
library(rgdal)
library(cowplot)

# Set your ggplot theme
mytheme <- theme_light(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "bottom", legend.justification = "right")
theme_set(mytheme)

# Load your datasets
tracklogs <- read.csv("./Data/Processed/IV_tracklogs_processed.csv")
wpts <- read.csv("./Data/Processed/IV_wpts_processed.csv")

#change the dates
tracklogs$date <- as.Date(tracklogs$date, format = "%Y-%m-%d")
wpts$date <- as.Date(wpts$date, format = "%Y-%m-%d")

```

# Rationale and Research Questions

## Elephant Movement in Gabon, Africa

@@

## Study Questions

@@

1. What is frequency and number of days traveled on each trail? 
2. What is the condition and use of the trail?

@@

# Dataset Information

@@

The waypoints dataset includes the waypoints taken by field teams when they started walking on an elephant trail or if the elephant trail changed characteristics. The data contained in this dataset includes date, time, latitude, longitude, trail characteristics (cmt), data on where the elephant trails started and stopped and when the field team got on or of a trail (name), and the symbol used in the GPS (sym). More specifically, the "name" field includes "start" or "stop" if the team got on or off a trail, "trstart" or "trstop" if the trail itself started or disappeared, "change" if a characteristic of the trail change, and "jct" if they came to a trail crossing. In the cmt field, there are two sets of information: a letter (P = poor, M = medium, H = heavy) representing how heavily trodden the path is and a number representing the width of the trail in cm. If the cmt section is blank then the details previously recorded remained the same. Data contained in this dataset was also recorded in decimal degrees WGS84 and the dates of data collection also range from February 2018 to May 2018.

__Dataset__  | __Variables__| __Range or Unique Values__| __Central Tendancies__
-------------|--------------|---------------------------|------------------------
IV_tracklogs |site          |                           |                                                              -       |date          |2018-02-05 - 2018-05-25    |
     -       |time          |                           |
     -       |latitude      |-0.3429768 - 0.6005945     |mean = 0.07083619; median = -0.1230693
     -       |longitude     |12.43589 - 12.81332        |mean = 12.60479; median = 12.60231
IV_wgts      |date          |2018-02-06 - 2018-05-25    |
     -       |time          |                           |
     -       |latitude      |-0.335521 - 0.599461       |mean = 0.2283088; median = 0.486579
     -       |longitude     |12.43817 - 12.80200        |mean = 12.66357; median = 12.72618
     -       |name          |                           |
     -       |sym           |                           |
     -       |cmt (letter)  | "P" "M" "W"               |
     -       |cmt (number)  | 35 - 75                   |mean = 50; median = 47.5

## Data Wrangling

@@

Because there are many observations on each day, we next combined the date and time columns to create a combined date_time column in each of the two datasets and arranged both datasets by ascending order of date_time. As stated earlier, the two datasets do not have exact matching date_time entires, so to assign cmt column values to the tracklogs dataset based on date_time, we transferred cmt values to the tracklogs dataset if the tracklogs date_time was within a window of the date_time column of the waypoints dataset using a for loop. We then filtered out all of the "NA" values of the resulting tracklogs dataset. We were told that if the "cmt" column is blank then the details previously recorded remained the same; Therefore, we filled in blanks with previously recorded values by first filling cells of empty strings with "NA" and then using the fill() command. Because the cmt column holds two pieces of information (a letter representing how heavily trodden the path is and a number representing the width of the trail), we separated these into two separate columns using the str_extract() command. Last, we assigned the "name" column values to the tracklogs dataset in the same way as we assigned the cmt values, using a for loop.

@@

# Exploratory Analysis

@@

__Use of Trail __            | __Total Amount__
--------------------         |---------------------------
Poor (P)                     | 27,218
Medium (M)                   | 13,666
Heavy (W)                    | 26

@@

```{r, echo=FALSE, fig.cap="Map of all trail use types on April 15, 2018.", fig.align="left", warning=FALSE}
##Insert code for map
##subset for each type of trail for 04-15-2018
Poor_trails <- tracklogs %>%
  filter(use_cha == "P")

Med_trail <- tracklogs %>% 
  filter(use_cha == "M")

Heavy_trail <- tracklogs %>%
  filter(use_cha == "W")

#filter by date
subset_poor <- Poor_trails %>%
  filter(date == "2018-04-15")

subset_med <- Med_trail %>%
  filter(date == "2018-04-15")

subset_heavy <- Heavy_trail %>%
  filter(date == "2018-04-15")

#CRS=WGS84 4326

#convert to sf and include crs
subset_poor_sf <- st_as_sf(subset_poor, crs = 4326, coords = c("lon", "lat"))
subset_med_sf <- st_as_sf(subset_med, crs = 4326, coords = c("lon", "lat"))
subset_heavy_sf <- st_as_sf(subset_heavy, crs = 4326, coords = c("lon", "lat"))

#Plot so each type of trail is a different color on 4- 15-18 so heavy is included 
Trail_map2 <- 
mapview(subset_poor_sf, color = "darkseagreen3", col.regions = "darkseagreen4") +
  mapview(subset_med_sf, color = "coral2", col.regions = "coral3") +
    mapview(subset_heavy_sf, color = "lightblue", col.regions = "darkblue")
Trail_map2
```

@@

# Analysis

@@

```{r, echo=FALSE, fig.cap="The poor trodden trails are more abundant earlier in the year and the medium trodden trails take over in late April-May timeframe.", fig.align="left", warning=FALSE}

##for caption: The poor trodden trails are more abundant earlier in the year and the medium trodden trails take over in late April-May timeframe.

freq_plot <- ggplot(tracklogs, aes(x=date, fill = use_cha)) + 
  geom_bar() + 
  scale_fill_manual(values = c("darkseagreen3", "deepskyblue3", "deeppink")) +
  scale_x_date(date_breaks ="10 days", date_labels = "%Y-%m-%d", 
          guide = guide_axis(angle = 45)) + 
  (theme(legend.position = "right", 
        legend.text = element_text(size = 12), legend.title = element_text(size = 12))) +
  labs( y = "Count", 
        x = "Date", 
        title = "Number of Trail Points per Day", 
        fill = "Use of Trail") +
  mytheme

print(freq_plot)

```

@@

```{r plots by width, echo=FALSE, fig.cap="On the x axis, the three letters represent how heavily trodden the path is:  (P = poor, M = medium, W = heavy) and the width of the trail on the y-axis in centimeters.", fig.align="left", warning=FALSE}
##for caption: On the x axis, the three letters represent how heavily trodden the path is:  (P = poor, M = medium, W = heavy) and the width of the trail on the y-axis in centimeters. It is evident the poor and medium trails are similar in width and the heavy trails are much wider.

width_plots <- ggplot(tracklogs, aes(x = use_cha, y = use_num)) +
  geom_boxplot(fill = "lightblue") + ggtitle("Trail type by width") + xlab("Type of Trail") + ylab("Width of trail (cm)") + mytheme

print(width_plots)

```

```{r, echo=FALSE, fig.cap="Post-hoc Tukey test for trail use.", fig.align="left", warning=FALSE}
mod3 <- lm(data = tracklogs, use_num ~ use_cha)
  plot(TukeyHSD(aov(mod3)))
```

@@

# Summary and Conclusions

@@

In our final dataset, which included only the data that were recorded on common dates between our two initial datasets, there are 27,218 poor trodden trail points, 13,666 medium trodden trail points, and 26 heavily trodden trail points. From our analyses on the widths of each type of trail, we found that the heavily trodden trail is significantly wider than the poor and medium trodden trails and the widths of the poor and medium trodden trails are not significantly different. While the heavily trodden trails are significantly wider than the poor and medium trodden trails, it is important to note that the heavily trodden trails have a much smaller number of recorded trail points, which occur in a compact geographic area compared to all of the other recorded points, and only on one day. Overall, we can conclude that the majority of recorded trail points are poorly trodden (67%) and medium trodden (33%), and very few are heavily trodden (<0.1%).

@@

>>>>

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
    \setmainfont[]{Times New Roman}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Analyzing Forest Elephant Trails in Gabon, Africa},
            pdfauthor={Elise Harrigan and Katie Krejsa},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=2.54cm]{geometry}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{longtable}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother

\title{Analyzing Forest Elephant Trails in Gabon, Africa}
\providecommand{\subtitle}[1]{}
\subtitle{\url{https://github.com/kmk85/HarriganKrejsa_ENV872_EDA_FinalProject}}
\author{Elise Harrigan and Katie Krejsa}
\date{}

\begin{document}
\maketitle

@@

\hypertarget{rationale-and-research-questions}{%
\section{Rationale and Research
Questions}\label{rationale-and-research-questions}}

\hypertarget{elephant-movement-in-gabon-africa}{%
\subsection{Elephant Movement in Gabon,
Africa}\label{elephant-movement-in-gabon-africa}}

@@

\hypertarget{study-questions}{%
\subsection{Study Questions}\label{study-questions}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  What is frequency and number of days traveled on each trail?
\item
  What is the condition and use of the trail?
\end{enumerate}

@@

\hypertarget{dataset-information}{%
\section{Dataset Information}\label{dataset-information}}

@@

The waypoints dataset includes the waypoints taken by field teams when
they started walking on an elephant trail or if the elephant trail
changed characteristics. The data contained in this dataset includes
date, time, latitude, longitude, trail characteristics (cmt), data on
where the elephant trails started and stopped and when the field team
got on or of a trail (name), and the symbol used in the GPS (sym). More
specifically, the ``name'' field includes ``start'' or ``stop'' if the
team got on or off a trail, ``trstart'' or ``trstop'' if the trail
itself started or disappeared, ``change'' if a characteristic of the
trail change, and ``jct'' if they came to a trail crossing. In the cmt
field, there are two sets of information: a letter (P = poor, M =
medium, H = heavy) representing how heavily trodden the path is and a
number representing the width of the trail in cm. If the cmt section is
blank then the details previously recorded remained the same. Data
contained in this dataset was also recorded in decimal degrees WGS84 and
the dates of data collection also range from February 2018 to May 2018.

\begin{longtable}[]{@{}llll@{}}
\toprule
\textbf{Dataset} & \textbf{Variables} & \textbf{Range or Unique Values}
& \textbf{Central Tendancies}\tabularnewline
\midrule
\endhead
IV\_tracklogs & site & & -\tabularnewline
- & time & &\tabularnewline
- & latitude & -0.3429768 - 0.6005945 & mean = 0.07083619; median =
-0.1230693\tabularnewline
- & longitude & 12.43589 - 12.81332 & mean = 12.60479; median =
12.60231\tabularnewline
IV\_wgts & date & 2018-02-06 - 2018-05-25 &\tabularnewline
- & time & &\tabularnewline
- & latitude & -0.335521 - 0.599461 & mean = 0.2283088; median =
0.486579\tabularnewline
- & longitude & 12.43817 - 12.80200 & mean = 12.66357; median =
12.72618\tabularnewline
- & name & &\tabularnewline
- & sym & &\tabularnewline
- & cmt (letter) & ``P'' ``M'' ``W'' &\tabularnewline
- & cmt (number) & 35 - 75 & mean = 50; median = 47.5\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{data-wrangling}{%
\subsection{Data Wrangling}\label{data-wrangling}}

@@

Because there are many observations on each day, we next combined the
date and time columns to create a combined date\_time column in each of
the two datasets and arranged both datasets by ascending order of
date\_time. As stated earlier, the two datasets do not have exact
matching date\_time entires, so to assign cmt column values to the
tracklogs dataset based on date\_time, we transferred cmt values to the
tracklogs dataset if the tracklogs date\_time was within a window of the
date\_time column of the waypoints dataset using a for loop. We then
filtered out all of the ``NA'' values of the resulting tracklogs
dataset. We were told that if the ``cmt'' column is blank then the
details previously recorded remained the same; Therefore, we filled in
blanks with previously recorded values by first filling cells of empty
strings with ``NA'' and then using the fill() command. Because the cmt
column holds two pieces of information (a letter representing how
heavily trodden the path is and a number representing the width of the
trail), we separated these into two separate columns using the
str\_extract() command. Last, we assigned the ``name'' column values to
the tracklogs dataset in the same way as we assigned the cmt values,
using a for loop.

@@

\hypertarget{exploratory-analysis}{%
\section{Exploratory Analysis}\label{exploratory-analysis}}

@@

\begin{longtable}[]{@{}ll@{}}
\toprule
\textbf{Use of Trail } & \textbf{Total Amount}\tabularnewline
\midrule
\endhead
Poor (P) & 27,218\tabularnewline
Medium (M) & 13,666\tabularnewline
Heavy (W) & 26\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{verbatim}
## PhantomJS not found. You can install it with webshot::install_phantomjs(). If it is installed, please make sure the phantomjs executable can be found via the PATH variable.
\end{verbatim}

@@

\hypertarget{analysis}{%
\section{Analysis}\label{analysis}}

@@

\begin{figure}

\includegraphics{Project_Template_files/figure-latex/unnamed-chunk-2-1} \hfill{}

\caption{The poor trodden trails are more abundant earlier in the year and the medium trodden trails take over in late April-May timeframe.}\label{fig:unnamed-chunk-2}
\end{figure}

@@

\begin{figure}

\includegraphics{Project_Template_files/figure-latex/plots by width-1} \hfill{}

\caption{On the x axis, the three letters represent how heavily trodden the path is:  (P = poor, M = medium, W = heavy) and the width of the trail on the y-axis in centimeters.}\label{fig:plots by width}
\end{figure}

\begin{figure}

\includegraphics{Project_Template_files/figure-latex/unnamed-chunk-3-1} \hfill{}

\caption{Post-hoc Tukey test for trail use.}\label{fig:unnamed-chunk-3}
\end{figure}

@@

\hypertarget{summary-and-conclusions}{%
\section{Summary and Conclusions}\label{summary-and-conclusions}}

@@

In our final dataset, which included only the data that were recorded on
common dates between our two initial datasets, there are 27,218 poor
trodden trail points, 13,666 medium trodden trail points, and 26 heavily
trodden trail points. From our analyses on the widths of each type of
trail, we found that the heavily trodden trail is significantly wider
than the poor and medium trodden trails and the widths of the poor and
medium trodden trails are not significantly different. While the heavily
trodden trails are significantly wider than the poor and medium trodden
trails, it is important to note that the heavily trodden trails have a
much smaller number of recorded trail points, which occur in a compact
geographic area compared to all of the other recorded points, and only
on one day. Overall, we can conclude that the majority of recorded trail
points are poorly trodden (67\%) and medium trodden (33\%), and very few
are heavily trodden (\textless{}0.1\%).

@@

\end{document}
